---
title: "Dictionaries and crosswalks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dictionaries and crosswalks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
```

The pipeline is driven by two kinds of reference data: **crosswalks** (curated
lookup tables shipped as package data) and **word lists** (character vectors shipped as package data used by
the regex-based cleaning steps).

## Crosswalks

The three crosswalks are package data (`title.xwalk`, `title.taxonomy`,
`status.codes`). Their source is the CSV tables in `data-raw/crosswalks/` of the
source repository: edit a table, then run `data-raw/crosswalks/build-crosswalks.R`
to rebuild the package data. (Until October 2026 the crosswalks were kept in a
Google Sheet; it is retired, and an export of every tab is archived in
`data-raw/crosswalks/archive/`.)

### Title standardization (`get_title_xwalk()`)

Maps every observed title **variant** to a canonical `title.standard`, plus a
`strata` / `strata.label` for board sub-roles. Used in step 07,
`standardize_titles()`.

```{r}
xwalk <- get_title_xwalk()
dim(xwalk)
head(xwalk, 8)
```

### Title taxonomy (`get_title_taxonomy()`)

Maps each canonical `title.standard` to its taxonomy dimensions: functional
`domain.category` / `domain.label`, the BLS SOC occupation codes, and its role.
A title is either an employee level or a board role:

- `emp.level`:
  - `CEO`: the top staff job.
  - `OFFICER`: C-suite, vice presidents, deputies.
  - `MANAGER`: directors of departments, managers.
  - `PROFESSIONAL`: non-managerial jobs often reported for their pay, such as
    physicians, lawyers, professors and coaches.
  - `STAFF`: everything else.
- `board.role`: `CHAIR`, `VICE CHAIR`, `SECRETARY`, `TREASURER` or `MEMBER`.

The legacy flags (`ceo`, `c.level`, `mgr`, `spec`, `board`, `pres`, ...) are
derived from these two columns. `dir.vp` is retired. Used in step 08,
`categorize_titles()`.

```{r}
tax <- get_title_taxonomy()
dim(tax)
head(tax[, c("title.standard", "domain.label", "soc.label",
             "emp.level", "board.role")], 8)
table(tax$emp.level)
table(tax$board.role)
```

### Status codes (`get_status_codes()`)

Maps status-qualifier **variants** (e.g. `ACTING`, `EMERITUS`, `INTERIM`) to a
standardized `status.qualifier` (e.g. `INTERIM`, `FORMER`). Used in step 06,
`gen_status_codes()`.

```{r}
sc <- get_status_codes()
dim(sc)
head(sc, 8)
```

## Word lists (package data)

These character vectors are lazy-loaded package datasets used by the regex cleaning
steps. Load any of them with `data()`.

| Dataset | Used in | Purpose |
|---------|---------|---------|
| `date.words` | `remove_dates()` (02) | month/date tokens to detect and strip |
| `number.words` | `gen_status_codes()` (06) | number words removed from titles |
| `likely.titles` | `standardize_conj()` (03) | recognizable titles for split detection |
| `likely.subjects` | `fix_of()` (05) | subjects that follow "of" in a title |
| `possible.titles` | `fix_of()` (05) | candidate title tokens |
| `former.words` | `gen_status_codes()` (06) | FORMER status variants |
| `future.words` | `gen_status_codes()` (06) | FUTURE status variants |
| `interim.words` | `gen_status_codes()` (06) | INTERIM status variants |
| `regional.words` | `gen_status_codes()` (06) | REGIONAL status variants |
| `possible.regional.list` | `gen_status_codes()` (06) | candidate regional tokens |

```{r}
data(number.words)
head(number.words, 12)

data(likely.titles)
head(likely.titles, 12)
```

## Demo data

| Dataset | Description |
|---------|-------------|
| `tinypartvii` | ~81k raw Part VII rows (10,000 orgs) in the modern schema, ready for `standardize_df()`. |
| `raw.titles` | A small vector of raw title strings for quick experiments. |

```{r}
data(tinypartvii)
str(tinypartvii[, 1:8], vec.len = 1)
```
