---
title: "Step 04 - Split multi-titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 04 - Split multi-titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
```

With separators normalized to `&`, **`split_titles()`** expands a row that holds
several roles into one row per role, duplicating the person's other data.

```
split_titles()
|
+- apply_misc_split_rules()  regex fixes for tricky edge cases
+- identify_split_num()      how many distinct titles does the string hold?
+- remove_first_split()      peel the text before the first "&"
```

## Counting and peeling

```{r}
identify_split_num("CFO & TREASURER & DIRECTOR")   # 3
remove_first_split("CFO & TREASURER & DIRECTOR")   # " TREASURER & DIRECTOR"
```

## In the pipeline

A single multi-title row becomes several rows, each carrying a `Num.Titles` index:

```{r}
data(tinypartvii)
set.seed(4)
d <- dplyr::sample_n(tinypartvii, 120)

df <- d |>
  standardize_df() |>
  remove_dates() |>
  standardize_conj() |>
  split_titles()

df |>
  dplyr::filter(grepl("&", TitleTxt3)) |>
  dplyr::select(F9_07_COMP_DTK_NAME_PERS, Num.Titles, TitleTxt3, TitleTxt4) |>
  head(10)
```

**Input:** `TitleTxt3`. **Output:** expanded rows with a single-title `TitleTxt4`
and a within-person `Num.Titles`. Continue with
[step 05](step-05-standardize-spelling.html).
