---
title: "Step 05 - Standardize spelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 05 - Standardize spelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
```

The dictionary-driven layer. **`standardize_spelling()`** applies `fix_spelling()`
to every title, which in turn calls ~50 small `fix_*` routines to expand
abbreviations and unify variants, then condenses unambiguous C-suite forms.

```
standardize_spelling()
+- fix_spelling()
   +- fix_vice(), fix_executive(), fix_director(), fix_president(), ...   (~50 fix_* routines)
   +- condense_abbreviations(), simplify_clevels()
   +- remove_trailing_conjunctions(), fix_of()
   +- spellcheck()   (hunspell)
```

## Examples

```{r}
fix_spelling("VP")                    # VICE PRESIDENT
fix_spelling("SR D MARKETING")        # SENIOR DIRECTOR OF MARKETING
fix_spelling("SECY")                  # SECRETARY
fix_spelling("TRTEE")                 # TRUSTEE
```

An implicit "of" is inserted before a recognized subject:

```{r}
fix_spelling("VICE PRESIDENT FINANCE")   # VICE PRESIDENT OF FINANCE
```

Unambiguous C-suite forms are condensed; ambiguous ones are left expanded.
(Note: EXECUTIVE DIRECTOR is normalized here but mapped to CEO later, by the
step 07 crosswalk.)

```{r}
fix_spelling("CHIEF EXEC OFFICER")        # CEO
fix_spelling("CHIEF FINANCIAL OFFICER")   # CFO
fix_spelling("CHIEF ADVANCEMENT OFFICER") # unchanged (no standard abbreviation)
```

Individual routines can be called directly — useful when refining rules:

```{r}
fix_vice(c("EVP", "SVP", "V.P."))
```

**Input:** `TitleTxt4`. **Output:** `TitleTxt5`, spelling-normalized. Continue with
[step 06](step-06-gen-status-codes.html).
