---
title: "Step 08 - Categorize titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 08 - Categorize titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
```

**`categorize_titles()`** joins the standardized titles to the taxonomy, engineers
organization-level features, and produces the finished, analysis-ready data frame.

```
categorize_titles()
|
+- merge with the taxonomy on title.standard  (domain, SOC codes, role flags)
+- add_features()        pay/hours ranks, leadership counts, per-EIN summaries
+- simplify_varnames()   rename to the tidy output schema; zero-fill ratio NAs
```

## Engineered features

`add_features()` computes, per organization (`OBJECTID`): counts of employees,
board members, and leaders (`num.ceos`, `num.clevel`, `num.dirvp`, `num.mgr`,
`num.spec`); officer counts (`num.pres`, `num.vp`, `num.treas`, `num.sec`); pay and
hours ranks and shares; and full/part-time headcounts — weighting people who hold
multiple titles so they are not double-counted.

```{r}
data(tinypartvii)
set.seed(8)
d <- dplyr::sample_n(tinypartvii, 250)

df <- d |>
  standardize_df() |> remove_dates() |> standardize_conj() |>
  split_titles() |> standardize_spelling() |> gen_status_codes() |>
  standardize_titles() |> categorize_titles()

df |>
  dplyr::select(org.name, dtk.name, title.standard, domain.category,
                pay.rank, tot.comp, num.ceos, num.board) |>
  dplyr::arrange(org.name, pay.rank) |>
  head(10)
```

## Where this leaves us

Every row is now a single standardized, categorized title enriched with
organization context. The remaining ambiguity (which "PRESIDENT" is the CEO?) is
the target of the forthcoming machine-learning relabeling step, which will consume
exactly these engineered features.

**Input:** `title.standard`. **Output:** the final analysis-ready data frame.
