---
title: "Step 06 - Generate status codes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 06 - Generate status codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
```

Titles often carry a temporal/temporary qualifier — FORMER, INTERIM, FUTURE,
EX OFFICIO, and so on. **`gen_status_codes()`** detects each qualifier, adds a
boolean flag, and either removes it or standardizes it in place.

```
gen_status_codes()
|
+- flag_and_remove()   flag SCODE.X and delete the qualifier (FORMER, INTERIM, FUTURE, ...)
+- flag_and_keep()     flag SCODE.X but keep a standardized form (EXOFFICIO, REGIONAL)
+- get_variants()      collapse a status code's variants into one regex
```

The variant-to-code mapping comes from the bundled status-code crosswalk (see the
[Dictionaries](dictionaries.html) article). Some qualifiers (EX OFFICIO, REGIONAL)
are kept because they change the meaning of the role.

## Examples

```{r}
sc <- get_googlesheets_status_codes()

# all variants that map to FORMER, as a single regex
get_variants("FORMER", sc)

x <- c("IMMEDIATE PAST CHAIR", "OUTGOING CEO", "ACTING DIRECTOR", "TREASURER")
remove_status(x, get_variants("FORMER",  sc))
remove_status(x, get_variants("INTERIM", sc))
```

## In the pipeline

```{r}
library(dplyr)
data(tinypartvii)
set.seed(6)
d <- dplyr::sample_n(tinypartvii, 150)

df <- d |>
  standardize_df() |> remove_dates() |> standardize_conj() |>
  split_titles() |> standardize_spelling() |> gen_status_codes()

df |>
  dplyr::filter(FORMER.X == 1 | INTERIM.X == 1) |>
  dplyr::select(TitleTxt5, TitleTxt6, FORMER.X, INTERIM.X) |>
  head(8)
```

**Input:** `TitleTxt5`. **Output:** `TitleTxt6` plus status flags (`FORMER.X`,
`INTERIM.X`, `FUTURE.X`, `EXOFFICIO.X`, `REGIONAL.X`, …). Continue with
[step 07](step-07-standardize-titles.html).
