---
title: "Step 05 - Standardize spelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 05 - Standardize spelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

The dictionary layer. **`standardize_spelling()`** expands abbreviations,
fixes common misspellings and truncations, and writes each kind of title one
way, so that the crosswalk in [step 07](step-07-standardize-titles.html) needs
one entry per title rather than one per spelling.

## Before and after

```{r, echo = FALSE}
x <- c("Exec. Dir.", "Sec", "Treas", "Asst. Secretary", "Dir of Development", "Dir. of Ops.",
       "Chief Exec Officer", "Chief Operating Officer", "Exec. V.P.", "SVP, Marketing",
       "VP Finance and Administration", "VP of HR", "Gen. Counsel", "Chairman of the Board",
       "Chairperson", "Board Pres.", "Mgr, IT", "Dir of Nursing")
out <- quiet(standardize_spelling(split_titles(standardize_conj(remove_dates(standardize_df(example_frame(x)))))))
out <- out[order(out$F9_07_COMP_DTK_NAME_PERS), ]
knitr::kable(data.frame(`before (TitleTxt4)` = out$TitleTxt4, `after (TitleTxt5)` = out$TitleTxt5,
                        check.names = FALSE), row.names = FALSE)
```

## What it does

```
standardize_spelling()
+- fix_spelling()
   +- fix_vice(), fix_executive(), fix_director(), fix_president(), ...   (about 50 fix_* routines)
   +- condense_abbreviations(), simplify_clevels()
   +- remove_trailing_conjunctions(), fix_of()
   +- spellcheck()   (hunspell)
```

* Each `fix_*` routine handles one word family: `fix_vice()` turns VP, V.P.,
  VICE-PRES and EVP into their full forms, `fix_director()` handles DIR, DIRS,
  DRCTR, and so on.
* `fix_of()` inserts the implicit "of" between a title and its subject:
  VICE PRESIDENT FINANCE becomes VICE PRESIDENT OF FINANCE.
* Unambiguous C-suite titles are condensed (CHIEF EXECUTIVE OFFICER becomes
  CEO, CHIEF FINANCIAL OFFICER becomes CFO). Titles without a standard
  abbreviation stay spelled out.
* `spellcheck()` suggests a dictionary spelling for a word no routine knows.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `possible.titles` | word list (package data) | title words that can precede an implicit "of" | `data(possible.titles)` |
| `likely.subjects` | word list (package data) | subjects that follow "of" (FINANCE, DEVELOPMENT, ...) | `data(likely.subjects)` |
| the `fix_*` routines | R functions | the abbreviation and spelling rules themselves | `?fix_spelling`, `?fix_vice`, ... |
| hunspell | R package (Imports) | the spell-check dictionary | `hunspell::dictionary("en_US")` |

```{r}
data(possible.titles)
head(possible.titles, 10)
data(likely.subjects)
head(likely.subjects, 15)
```

## Try it

```{r}
fix_spelling("VP")                        # VICE PRESIDENT
fix_spelling("SR D MARKETING")            # SENIOR DIRECTOR OF MARKETING
fix_spelling("SECY")                      # SECRETARY
fix_spelling("VICE PRESIDENT FINANCE")    # VICE PRESIDENT OF FINANCE
fix_spelling("CHIEF EXEC OFFICER")        # CEO
fix_spelling("CHIEF ADVANCEMENT OFFICER") # unchanged: no standard abbreviation
fix_vice(c("EVP", "SVP", "V.P."))
```

**Input:** `TitleTxt4`. **Output:** `TitleTxt5`, with standard spelling.
Continue with [step 06](step-06-gen-status-codes.html).
