---
title: "Step 06 - Generate status codes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 06 - Generate status codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

Many titles carry a qualifier about the person's status in the role: FORMER,
INTERIM, PRESIDENT-ELECT, EX OFFICIO, AT LARGE, CO-CHAIR. The qualifier
matters for analysis (an interim CEO, a past president) but would split one
title into many crosswalk entries. **`gen_status_codes()`** detects each
qualifier, records it as a 0/1 flag, and removes it from the title, or, for a
few that change the role, keeps a standard form.

## Before and after

```{r, echo = FALSE}
x <- c("Former CFO", "Acting CFO", "Interim Executive Director", "President-Elect", "Past President",
       "Outgoing Treasurer", "Emeritus Trustee", "Board Member (thru 6/2023)", "Trustee, Ex-Officio",
       "Member at Large", "Co-Chair", "1st Vice President", "Regional Director",
       "Founder & Executive Director", "Treasurer")
out <- quiet(gen_status_codes(standardize_spelling(split_titles(standardize_conj(remove_dates(
  standardize_df(example_frame(x))))))))
out <- out[order(out$F9_07_COMP_DTK_NAME_PERS, out$Num.Titles), ]
fl <- c("FORMER.X", "INTERIM.X", "FUTURE.X", "OUTGOING.X", "PARTIAL.X", "EXOFFICIO.X", "AT.LARGE.X",
        "CO.X", "REGIONAL.X", "FOUNDER.X", "SCHED.O.X")
flags <- apply(out[, fl], 1, function(r) paste(fl[r %in% 1], collapse = ", "))
knitr::kable(data.frame(`before (TitleTxt5)` = out$TitleTxt5, `after (TitleTxt6)` = out$TitleTxt6,
                        `flags set` = flags, check.names = FALSE), row.names = FALSE)
```

A person's flags apply to all of their titles: in "FOUNDER & EXECUTIVE
DIRECTOR", both rows carry `FOUNDER.X`.

## What it does

```
gen_status_codes()
|
+- remove number words and ordinals (FIRST, SECOND, ...)
+- flag_and_keep()     flag, write one standard form:  EXOFFICIO, REGIONAL
+- flag_and_remove()   flag, delete the qualifier:     FORMER, FOUNDER, FUTURE, INTERIM,
|                                                     OUTGOING, PARTIAL, SCHED O, AT LARGE
+- get_variants()      all the spellings of one status, as one regular expression
+- share flags across a person's rows
+- clean up            drop a trailing EXOFFICIO, a CO- prefix (flag CO.X), NON-VOTING,
                       stray parentheses and punctuation
```

* REGIONAL stays in the title, because a regional director is a different
  job. EXOFFICIO stays only when it is the whole title (EXOFFICIO BOARD
  MEMBER).
* `PARTIAL` marks a partial year: a date in the title, or an outgoing, future
  or part-year status.
* A title with both FORMER and FUTURE ("from ... until") keeps FORMER only.

The flags become `former.x`, `interim.x`, `future.x`, `outgoing.x`,
`partial.x`, `ex.officio.x`, `at.large.x`, `co.x`, `region.x`, `founder.x`,
`sched.o.x` and `date.x` in the final output. The
[Taxonomy](taxonomy.html#status-flags) article defines each one.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `status.codes` | crosswalk (package data) | maps each qualifier spelling (`status.variant`) to its status (`status.qualifier`) | `get_status_codes()` or `data(status.codes)` |
| `number.words` | word list (package data) | number words and ordinals to remove | `data(number.words)` |

The crosswalk's source is `data-raw/crosswalks/status-codes.csv` in the source
repository.

```{r}
sc <- get_status_codes()
dim(sc)
table(sc$status.qualifier)
head(sc[sc$status.qualifier == "INTERIM", ], 10)
```

## Try it

```{r}
get_variants("FORMER", sc)    # all the spellings of FORMER, as one regex

x <- c("IMMEDIATE PAST CHAIR", "OUTGOING CEO", "ACTING DIRECTOR", "TREASURER")
remove_status(x, get_variants("FORMER",  sc))
remove_status(x, get_variants("INTERIM", sc))
```

**Input:** `TitleTxt5`. **Output:** `TitleTxt6`, without status qualifiers,
and one flag per status. Continue with
[step 07](step-07-standardize-titles.html).
