---
title: "Step 09 - Resolve roles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 09 - Resolve roles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

The crosswalk maps each title to one role, but some titles mean different
things in different organizations:

- A **PRESIDENT** is the paid chief executive in one organization and a
  volunteer board chair in the next.
- A **VICE PRESIDENT** may sit on the board or on staff.
- A **DIRECTOR** may be a board member or a department head.

**`conditional_logic()`** (a wrapper for `resolve_roles()`) settles these with
what the filer reported about each person: the Part VII checkboxes (trustee,
officer, key employee, highly compensated), pay and hours. It also names the
chief executive when no title does, and decides that a filing has none when
the board runs the organization.

## Before and after

Four example filings, built by hand. "Before" is the title-only answer from
steps 07-08. "After" is step 09's answer for each person.

```{r, include = FALSE}
# one hand-built filing: a row per person
filing <- function(id, form, people) {
  d <- titleclassifier::tinypartvii[rep(1, nrow(people)), ]
  d$OBJECTID <- id; d$RETURN_TYPE <- form
  d$F9_07_COMP_DTK_TITLE <- people$title
  d$F9_07_COMP_DTK_NAME_PERS <- paste(id, seq_len(nrow(people)))
  d$F9_07_COMP_DTK_AVE_HOUR_WEEK <- as.character(people$hours)
  d$F9_07_COMP_DTK_AVE_HOUR_WEEK_RL <- "0"
  d$F9_07_COMP_DTK_COMP_ORG <- as.character(people$pay)
  d$F9_07_COMP_DTK_COMP_RLTD <- "0"; d$F9_07_COMP_DTK_COMP_OTH <- "0"; d$F9_07_COMP_DTK_EMPL_BEN <- "0"
  box <- function(b) if (form == "990EZ") "" else ifelse(grepl(b, people$boxes), "X", "")
  d$F9_07_COMP_DTK_POS_INDIV_TRUST_X <- box("T"); d$F9_07_COMP_DTK_POS_INST_TRUST_X <- ""
  d$F9_07_COMP_DTK_POS_OFF_X <- box("O"); d$F9_07_COMP_DTK_POS_KEY_EMPL_X <- box("K")
  d$F9_07_COMP_DTK_POS_HIGH_COMP_X <- box("H"); d$F9_07_COMP_DTK_POS_FORMER_X <- ""
  rownames(d) <- NULL
  d
}
p <- function(...) data.frame(..., stringsAsFactors = FALSE)
A <- filing("A large organization (990)", "990", p(
  title = c("President & CEO", "Chair of the Board", "Treasurer", "Director", "Chief Financial Officer",
            "VP of Development", "Director of Nursing", "Staff Physician"),
  pay   = c(410000, 0, 0, 0, 260000, 190000, 175000, 320000),
  hours = c(55, 5, 3, 2, 50, 45, 40, 40),
  boxes = c("O", "TO", "TO", "T", "O", "K", "H", "H")))
B <- filing("B paid president, no CEO title (990)", "990", p(
  title = c("President", "Vice President", "Secretary", "Program Director"),
  pay   = c(85000, 0, 0, 42000), hours = c(40, 2, 2, 40), boxes = c("TO", "T", "TO", "")))
C <- filing("C run by its board (990)", "990", p(
  title = c("President", "Secretary", "Treasurer", "Bookkeeper"),
  pay   = c(6000, 0, 0, 18000), hours = c(5, 2, 2, 10), boxes = c("TO", "TO", "TO", "")))
D <- filing("D small organization (990-EZ)", "990EZ", p(
  title = c("President", "Administrator", "Director of Operations", "Secretary", "Treasurer"),
  pay   = c(38000, 30000, 0, 0, 0), hours = c(40, 40, 10, 2, 2), boxes = ""))
steps08 <- function(d) d |> standardize_df() |> remove_dates() |> standardize_conj() |> split_titles() |>
  standardize_spelling() |> gen_status_codes() |> standardize_titles() |> categorize_titles()
ex  <- quiet(steps08(rbind(A, B, C, D)))
res <- quiet(conditional_logic(ex))
res$rules.only <- quiet(resolve_roles(ex, use_model = FALSE))$role.final
show <- function(id) {
  r <- res[res$object.id == id & res$role.primary, ]
  r <- r[order(as.integer(sub(".* ", "", r$dtk.name))), ]
  if (r$formtype[1] == "990EZ") r$role.method <- paste0(r$role.method, " (rules alone: ", r$rules.only, ")")
  boxes <- with(r, paste0(ifelse(dtk.indiv.trustee.x %in% 1, "T", ""), ifelse(dtk.officer.x %in% 1, "O", ""),
                          ifelse(dtk.key.empl.x %in% 1, "K", ""), ifelse(dtk.high.comp.x %in% 1, "H", "")))
  before <- ifelse(nzchar(r$board.role), paste("BOARD", r$board.role), r$emp.level)
  after  <- ifelse(r$role.final == "BOARD", paste("BOARD", r$role.board),
            ifelse(r$role.final == "CEO", paste0("CEO (", r$role.ceo, ")"), r$role.final))
  knitr::kable(data.frame(title = r$title.raw, pay = format(r$tot.comp, big.mark = ","), hours = r$tot.hours,
                          boxes = if (r$formtype[1] == "990EZ") "none on a 990-EZ" else boxes,
                          `before: title only` = before, `after: step 09` = after,
                          method = r$role.method, check.names = FALSE), row.names = FALSE,
               caption = sprintf("%s. org.leadership: %s", id, r$org.leadership[1]))
}
```

Boxes: **T** trustee, **O** officer, **K** key employee, **H** highly
compensated.

```{r, echo = FALSE}
show("A large organization (990)")
```

The checkboxes settle most of filing A. The CEO is `designated` by a
CEO-level title in an officer position. The board titles with the trustee box
are board seats. The paid staff keep the level their title implies.

```{r, echo = FALSE}
show("B paid president, no CEO title (990)")
```

No title in filing B names a CEO, but the president is paid $85,000 for 40
hours a week and ticks the officer box: the president is the `imputed` CEO.
The unpaid vice president is a board seat (VICE CHAIR), not a staff officer.

```{r, echo = FALSE}
show("C run by its board (990)")
```

Filing C pays its president a $6,000 stipend for 5 hours a week. A board title
held under 10 hours a week for under $25,000 is a board seat, even with the
officer box. No paid person qualifies as a leader (the bookkeeper is staff),
so the filing is `board_governed`: no CEO is named.

```{r, echo = FALSE}
show("D small organization (990-EZ)")
```

A 990-EZ has no checkboxes. When the xgboost package is installed, roles on a
990-EZ come from the no-checkbox model, and `role.method` says `model`.
Without xgboost, or with `resolve_roles(use_model = FALSE)`, the rules decide.
Here they disagree on two people:
- **Director of Operations.** The rules read the title as a department head
  (MANAGER). The model reads an unpaid person working 10 hours a week in a
  small filing as a volunteer board member.
- **Administrator.** The model makes the second paid full-timer an officer
  rather than a manager.

[Rules and models](classifiers.html) explains the difference.

## What it does

```
conditional_logic()  ->  resolve_roles()
|
+- 1. position, per row       trustee box = board; officer box = officer; both = board_officer;
|                             key-employee or high-comp box = staff; a paid CEO- or
|                             officer-level title counts as an officer without the box
+- 2. one role per person     a person with split titles (PRESIDENT & TREASURER) gets
|                             the most senior position; role.primary marks one row
+- 3. leadership, per filing  a CEO-level title in an officer position = designated CEO;
|                             else the highest-paid plausible leader = imputed CEO;
|                             else board_governed (paid people, no leader) or no_paid
+- 4. final role              CEO, OFFICER, MANAGER, PROFESSIONAL, STAFF or BOARD,
|                             and for BOARD the board role
+- 5. 990-EZ                  the no-checkbox model replaces the rules' answer
```

Three thresholds keep token pay from making a leader:

* **Imputed CEO.** A leader with no CEO title must work at least 15 hours a
  week or earn at least $25,000, and have the officer box or a leadership
  title.
* **Board chair pattern.** A paid board chair with the trustee box who works
  under 25 hours a week is not imputed as CEO.
* **Stipend rule.** A board title held under 10 hours a week for under
  $25,000 is a board seat, with or without the officer box. On a 990-EZ the
  same holds for a board title paid a small fee.

## Columns added

| Column | Values |
|---|---|
| `role.final` | CEO, OFFICER, MANAGER, PROFESSIONAL, STAFF, BOARD |
| `role.board` | for BOARD: CHAIR, VICE CHAIR, SECRETARY, TREASURER, MEMBER |
| `role.ceo` | for CEO: designated, co-ceo, interim, imputed, model |
| `role.position` | board, officer, board_officer, staff, unknown |
| `role.source` | what settled the position: checkbox, title+pay, title |
| `role.method` | rules, or model (990-EZ with xgboost installed) |
| `role.primary` | TRUE on one row per person (count people with it) |
| `org.leadership` | per filing: designated, imputed, model, board_governed, no_paid |

The crosswalk's own `emp.level` and `board.role` are kept: `role.final` is the
resolved answer, the crosswalk columns the title-only default.
[Taxonomy](taxonomy.html#roles) defines each value.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `emp.level`, `board.role` | columns of `title.taxonomy` (from step 08) | the title-only role each rule starts from | `get_title_taxonomy()` |
| the role model | xgboost model file, `inst/extdata/role-model/` | roles on 990-EZ filings | `predict_roles()`; features from `role_model_features()` |

```{r}
system.file("extdata", "role-model", package = "titleclassifier") |> list.files()
```

```{r, eval = requireNamespace("xgboost", quietly = TRUE)}
# the model's input: one row per person-title, no checkboxes
X <- role_model_features(ex[ex$object.id == "D small organization (990-EZ)", ])
dim(X)
head(colnames(X), 12)
predict_roles(ex[ex$object.id == "D small organization (990-EZ)", ])
```

## Try it on the demo data

```{r}
data(tinypartvii)
set.seed(8)
d <- dplyr::sample_n(tinypartvii, 1500)
df <- d |>
  standardize_df() |> remove_dates() |> standardize_conj() |>
  split_titles() |> standardize_spelling() |> gen_status_codes() |>
  standardize_titles() |> categorize_titles() |> conditional_logic()

people <- dplyr::filter(df, role.primary)
table(people$role.final)
table(people$formtype, people$role.method)
```

Titles the crosswalk files as board seats but that resolve otherwise, here a
paid board president who runs the organization:

```{r}
people |>
  dplyr::filter(title.standard == "BOARD PRESIDENT", role.final != "BOARD") |>
  dplyr::select(org.name, title.raw, tot.comp, tot.hours, role.final, role.ceo) |>
  head(8)
```

How each filing is led:

```{r}
df |>
  dplyr::distinct(object.id, formtype, org.leadership) |>
  dplyr::count(formtype, org.leadership)
```

**Input:** the output of `categorize_titles()`. **Output:** the same rows with
the role columns added. See [Performance](performance.html) for how often the
roles agree with hand labels.
