---
title: "Step 02 - Remove dates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 02 - Remove dates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

Filers often add the dates a person held a position to the title.
**`remove_dates()`** detects them, sets a `DATE.X` flag, and strips the date.
The words left behind (THRU, UNTIL, SINCE) are read as status qualifiers in
[step 06](step-06-gen-status-codes.html).

## Before and after

```{r, echo = FALSE}
x <- c("Board Member (thru 6/2023)", "Treasurer until Jan 2023", "Trustee (Through 12/2022)",
       "Director to 5/22", "President 2021-2023", "Secretary (resigned 3/16/23)",
       "1st Vice President", "2nd Vice Chair", "Executive Director")
out <- quiet(remove_dates(standardize_df(example_frame(x))))
knitr::kable(data.frame(`before (step 01)` = out$F9_07_COMP_DTK_TITLE,
                        `after (TitleTxt2)` = out$TitleTxt2,
                        DATE.X = out$DATE.X, check.names = FALSE))
```

## What it does

```
remove_dates()
|
+- convert_ordinal()  1ST/2ND/3RD -> FIRST/SECOND/THIRD
+- has_date()         does the string contain a date?        -> DATE.X
+- remove_date()      strip 'YY, YY-YY, mm/dd/yyyy, month names, trailing numbers
```

Ordinals are spelled out first, so "1ST VICE PRESIDENT" is not read as a date.
The ordinal itself (FIRST, SECOND) is removed later, in step 06.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `date.words` | word list (package data) | month names and abbreviations that mark a date | `data(date.words)` |

```{r}
data(date.words)
head(date.words, 15)
```

## Try it

```{r}
x <- c("TREASURER (ENDED 3/16/23)", "DIRECTOR THRU 12/23",
       "SECRETARY UNTIL JAN 2023", "PRESIDENT")
data.frame(title = x, has_date = has_date(x), cleaned = remove_date(x))

convert_ordinal(c("1ST VICE PRESIDENT", "2ND LIEUTENANT GOVERNOR"))
```

**Input:** the cleaned title from step 01. **Output:** `TitleTxt2`, the title
without dates, and the `DATE.X` flag. Continue with
[step 03](step-03-standardize-conj.html).
