---
title: "Step 07 - Standardize titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 07 - Standardize titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

After steps 01-06 a title is clean but still one of thousands of forms.
**`standardize_titles()`** maps each one to a canonical `title.standard`, one
of a few hundred titles that the taxonomy in
[step 08](step-08-categorize-titles.html) describes. It works in three passes:

1. **C-suite fixes** (`basic_csuite_fixes()`): a few titles are rewritten
   using pay and the officer box before the lookup. A paid CHANCELLOR is the
   CEO and an unpaid one the board president; a finance title held by a paid,
   full-time officer is the CFO.
2. **Exact match**: the title is looked up in the title-standardization
   crosswalk, which lists every known variant. `title.match` is `exact`.
3. **Pattern match**: a title the crosswalk does not list falls back to a
   learned rule for its longest known head. "MANAGER IT" ends in no known
   title but starts with MANAGER, which has a rule. `title.match` is
   `pattern`. Titles whose head depends on context (DIRECTOR, PRESIDENT) have
   no rule, and a title with no rule stays uncoded (`title.match` is NA).

## Before and after

```{r, echo = FALSE}
x <- c("Exec. Dir.", "Chief Exec Officer", "Head of School", "Pastor", "Director", "Dir",
       "Chairman of the Board", "Chairperson", "Co-Chair", "1st Vice President", "Vice Chair",
       "Treasurer", "Asst. Treasurer", "SVP, Marketing", "Exec. V.P.", "Advisory Board Member",
       "Regional Director", "Finance Committee Member", "Staff Physician", "Head Football Coach",
       "Camp Director", "Mgr, IT", "Hospitalist Physician",
       "Senior Rabbi", "Campus Advocate")
out <- quiet(standardize_titles(gen_status_codes(standardize_spelling(split_titles(standardize_conj(
  remove_dates(standardize_df(example_frame(x)))))))))
out <- out[order(out$F9_07_COMP_DTK_NAME_PERS), ]
knitr::kable(data.frame(`before (TitleTxt6)` = out$TitleTxt6, title.standard = out$title.standard,
                        title.match = out$title.match, strata = out$strata,
                        strata.label = out$strata.label, check.names = FALSE), row.names = FALSE)
```

Things to notice:

* Many variants collapse onto one standard: DIRECTOR and DIR become BOARD
  MEMBER, CHAIR and CHAIR OF BOARD become BOARD PRESIDENT.
* A plain DIRECTOR is a board seat. A director *of something* (CAMP
  DIRECTOR, DIRECTOR OF DEVELOPMENT) is a staff job.
* The crosswalk drops rank words such as SENIOR, EXECUTIVE and ASSISTANT from
  the standard and keeps them in `strata` and `strata.label`: SENIOR VICE
  PRESIDENT OF MARKETING is a VICE PRESIDENT OF MARKETING at the `superior` /
  `senior` stratum.
* The standard is the title-only answer. A PRESIDENT may be the paid chief
  executive or the volunteer board chair. Step 07 files it as BOARD
  PRESIDENT, and [step 09](step-09-conditional-logic.html) decides from pay,
  hours and the checkboxes.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `title.xwalk` | crosswalk (package data) | every known `title.variant` and its `title.standard`, `strata` and `strata.label` | `get_title_xwalk()` or `data(title.xwalk)` |
| `title.patterns` | rules (package data) | head rules for titles the crosswalk does not list | `get_title_patterns()` or `data(title.patterns)` |

Their sources are `data-raw/crosswalks/title-standardization.csv` and
`title-patterns.csv` in the source repository.

```{r}
xwalk <- get_title_xwalk()
dim(xwalk)
length(unique(xwalk$title.standard))

# all the variants that map to one standard
head(xwalk$title.variant[xwalk$title.standard == "CHIEF FINANCIAL OFFICER"], 15)

# the strata
table(xwalk$strata[xwalk$strata != ""])
```

Each pattern rule names a head word or phrase, where it sits (`prefix` or
`suffix`), the standard it maps to, and how well it did on titles the
crosswalk does list (`support` titles, `precision` the share it got right).
`majority` rules map to the most common standard among titles with that head;
`level` rules keep the head's own standard.

```{r}
pat <- get_title_patterns()
dim(pat)
head(pat[order(-pat$support), ], 10)
```

## Try it

```{r}
match_title_patterns(c("MANAGER IT", "HOSPITALIST PHYSICIAN", "SENIOR RABBI"))
title_heads(c("FINANCE COMMITTEE MEMBER", "VICE PRESIDENT OF CLINICAL INTEGRATION"),
            xwalk$title.variant)
```

**Input:** `TitleTxt6`. **Output:** `TitleTxt7` (after the C-suite fixes),
`title.standard`, `title.match`, `strata` and `strata.label`. Continue with
[step 08](step-08-categorize-titles.html).
