---
title: "Step 03 - Standardize conjunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 03 - Standardize conjunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
```

People list multiple roles with no standard punctuation ("CEO AND BOARD PRESIDENT",
"VP OF FINANCE AND ADMINISTRATION"). **`standardize_conj()`** normalizes the
connectors so that step 04 can split reliably, converting genuine separators into a
single `&` while leaving compound titles intact.

```
standardize_conj()
|
+- standardize_and()        AND -> & when both sides are valid titles
+- standardize_to()         parenthetical / trailing TO -> UNTIL (dates)
+- standardize_of()         AS OF -> SINCE; drop dangling OF; FOR -> OF
+- standardize_comma()      comma as separator / "of" / redundant
+- standardize_slash()      slash as separator / "of" / redundant
+- standardize_and()        re-applied to catch newly created cases
+- standardize_separator()  ";" and "\\" -> &
```

## The trick: separator vs. compound

The `and`/`&` decision checks whether both sides look like real titles:

```{r}
standardize_and("CEO AND BOARD PRESIDENT")        # separator -> &
standardize_and("VP OF FINANCE AND ADMINISTRATION") # compound  -> unchanged
```

Commas and slashes are disambiguated the same way — separator, "of", or filler:

```{r}
standardize_comma("SECRETARY, TREASURER")   # separator
standardize_comma("VP, MARKETING")          # "of"
standardize_slash("SECRETARY/TREASURER")    # separator
```

"TO" and "OF" at the end of a title usually belonged to a (removed) date phrase:

```{r}
standardize_to("DIRECTOR TO")               # -> "DIRECTOR UNTIL"
standardize_of("VICE PRESIDENT AS OF")      # -> "VICE PRESIDENT SINCE"
```

**Input:** `TitleTxt2`. **Output:** `TitleTxt3` with `&` as the single title
separator. Continue with [step 04](step-04-split-titles.html).
