---
title: "Step 01 - Standardize the data frame"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 01 - Standardize the data frame}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

**`standardize_df()`** turns a raw Part VII compensation table into the form
the rest of the pipeline expects: upper-case titles without periods, numeric
hours and pay, and 0/1 checkboxes.

## Before and after

Rows of the demo data, as filed and after step 01:

```{r, echo = FALSE}
data(tinypartvii)
raw <- tinypartvii |>
  filter(grepl("[a-z]", F9_07_COMP_DTK_TITLE), grepl("[.,/&]", F9_07_COMP_DTK_TITLE)) |>
  slice(c(2, 5, 9, 14, 20, 27, 35, 44))
out <- quiet(standardize_df(raw))
num <- function(x) { x <- suppressWarnings(as.numeric(x)); x[is.na(x)] <- 0; x }
knitr::kable(data.frame(
  form                  = raw$RETURN_TYPE,
  `title, filed`        = raw$F9_07_COMP_DTK_TITLE,
  `title, step 01`      = out$F9_07_COMP_DTK_TITLE,
  `hours, filed`        = raw$F9_07_COMP_DTK_AVE_HOUR_WEEK,
  TOT.HOURS             = out$TOT.HOURS,
  `pay, org`            = raw$F9_07_COMP_DTK_COMP_ORG,
  `other pay + benefits` = num(raw$F9_07_COMP_DTK_COMP_OTH) + num(raw$F9_07_COMP_DTK_EMPL_BEN),
  TOT.COMP              = out$TOT.COMP,
  `officer box, filed`  = raw$F9_07_COMP_DTK_POS_OFF_X,
  `officer box, step 01` = out$F9_07_COMP_DTK_POS_OFF_X,
  check.names = FALSE))
```

## What it does

```
standardize_df()
|
+- check_names()   add any missing compensation or checkbox column (as NA)
+- pre_clean()     upper-case the title and strip periods
+- to_numeric()    hours and the pay components as numbers, blanks as 0
+- to_boole()      checkboxes: "X" -> 1, blank -> 0; NA on a 990-EZ, which has no boxes
```

* Keeps the filed title as `TITLE_RAW` and gives each row a `PERSONID`.
* Sums hours at the organization and at related organizations into
  `TOT.HOURS`.
* Sums pay from the organization and from related organizations, other
  compensation and benefits into `TOT.COMP`.
* Converts the trustee, officer, key-employee, highly-compensated and former
  checkboxes to 1 or 0. A 990-EZ has no checkboxes, so on a 990-EZ they are NA.

## Assets

Step 01 uses no crosswalk or word list. It needs the Part VII column names of
the NCCS efile schema (`F9_07_COMP_DTK_*`). See the [Workflow](workflow.html)
article for fetching the tables. `tinypartvii` is a demo slice in that schema:

```{r}
data(tinypartvii)
dim(tinypartvii)
grep("^F9_07", names(tinypartvii), value = TRUE)
```

## Try it

```{r}
pre_clean(c("Vice President.", "sec/treas.", "Exec. Dir."))
```

```{r}
d  <- dplyr::sample_n(tinypartvii, 50)
df <- standardize_df(d)
df |>
  dplyr::select(TITLE_RAW, F9_07_COMP_DTK_TITLE, TOT.HOURS, TOT.COMP) |>
  head(6)
```

**Input:** a raw Part VII data frame (NCCS efile schema).
**Output:** the same rows with cleaned title text, numeric hours and pay,
0/1 checkboxes and a `PERSONID`. Continue with
[step 02](step-02-remove-dates.html).
