---
title: "Step 01 - Standardize the data frame"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 01 - Standardize the data frame}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
```

**`standardize_df()`** cleans the raw Part VII compensation table into a numeric-
friendly, consistently-named form that the rest of the pipeline expects.

```
standardize_df()
|
+- check_names()   ensure expected compensation/checkbox columns exist (add NA if missing)
+- pre_clean()     uppercase titles, strip periods         -> applied to the title field
+- to_numeric()    coerce hours/compensation to numeric, blanks -> 0
+- to_boole()      normalize "X"/"YES"/blank checkboxes to 1/0 (NA on 990EZ)
```

## What it does

* Adds a per-row `PERSONID` hash and preserves the raw title as `TITLE_RAW`.
* Uppercases titles and strips punctuation via `pre_clean()`.
* Converts hours and the four compensation components to numeric, summing
  `TOT.HOURS` and `TOT.COMP`.
* Normalizes the trustee/officer/key-employee/high-comp/former checkboxes to 0/1.

## Example

```{r}
pre_clean(c("Vice President.", "sec/treas."))
```

```{r}
data(tinypartvii)
set.seed(1)
d  <- dplyr::sample_n(tinypartvii, 50)
df <- standardize_df(d)

df |>
  dplyr::select(TITLE_RAW, F9_07_COMP_DTK_TITLE, TOT.HOURS, TOT.COMP) |>
  head(8)
```

**Input:** a raw Part VII data frame (modern NCCS schema).
**Output:** the same rows with cleaned title text, numeric money/hours, boolean
checkboxes, and a `PERSONID`. Continue with [step 02](step-02-remove-dates.html).
