---
title: "Step 03 - Standardize conjunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 03 - Standardize conjunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

People list several roles in one title with no standard punctuation: "CEO AND
BOARD PRESIDENT", "SEC/TREAS", "BOARD MEMBER; TREASURER". Other titles use the
same words inside a single role: "VP OF FINANCE AND ADMINISTRATION".
**`standardize_conj()`** turns every connector that separates two titles into
a single `&`, so that [step 04](step-04-split-titles.html) can split on it, and
leaves compound titles whole.

## Before and after

```{r, echo = FALSE}
x <- c("Sec/Treas", "Secretary, Treasurer", "Board Member; Treasurer", "Chief Exec Officer / Founder",
       "President and CEO", "VP Finance and Administration", "SVP, Marketing", "VP, Marketing",
       "Chairman of the Board", "Director to 5/22", "Trustee, Ex-Officio", "Mgr, IT")
out <- quiet(standardize_conj(remove_dates(standardize_df(example_frame(x)))))
knitr::kable(data.frame(`before (TitleTxt2)` = out$TitleTxt2, `after (TitleTxt3)` = out$TitleTxt3,
                        check.names = FALSE))
```

## What it does

```
standardize_conj()
|
+- standardize_and()        AND -> & when both sides are titles
+- standardize_to()         a trailing TO -> UNTIL (left by a removed date)
+- standardize_of()         AS OF -> SINCE; drop a dangling OF; FOR -> OF
+- standardize_comma()      a comma as separator, as "of", or as filler
+- standardize_slash()      a slash as separator, as "of", or as filler
+- standardize_and()        again, for cases the steps above created
+- standardize_separator()  ";" and "\" -> &
```

The decision for AND, a comma or a slash is the same: if the words on both
sides are recognizable titles, the connector separates two roles and becomes
`&`. If the words after it are a subject (FINANCE, MARKETING), it means "of".
Otherwise it is dropped.

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `likely.titles` | word list (package data) | titles that, found on both sides of a connector, make it a separator | `data(likely.titles)` |

```{r}
data(likely.titles)
length(likely.titles)
head(likely.titles, 20)
```

## Try it

```{r}
standardize_and("CEO AND BOARD PRESIDENT")           # separator -> &
standardize_and("VP OF FINANCE AND ADMINISTRATION")  # compound  -> unchanged
standardize_comma("SECRETARY, TREASURER")            # separator
standardize_comma("VP, MARKETING")                   # "of"
standardize_slash("SECRETARY/TREASURER")             # separator
standardize_to("DIRECTOR TO")                        # -> DIRECTOR UNTIL
standardize_of("VICE PRESIDENT AS OF")               # -> VICE PRESIDENT SINCE
```

**Input:** `TitleTxt2`. **Output:** `TitleTxt3`, with `&` as the only
separator between titles. Continue with [step 04](step-04-split-titles.html).
