---
title: "Step 04 - Split multi-titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 04 - Split multi-titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

With the separators reduced to `&`, **`split_titles()`** turns a row that
holds several roles into one row per role. The person's name, pay, hours and
checkboxes repeat on each row; `Num.Titles` numbers the rows. A few common
pairs are collapsed to one title rather than split: "PRESIDENT & CEO" is a CEO.

## Before and after

```{r, echo = FALSE}
x <- c("Sec/Treas", "Board Member; Treasurer", "Chief Exec Officer / Founder", "EVP & COO",
       "President & CEO", "Trustee & Treasurer", "Founder & Executive Director",
       "VP Finance and Administration", "Executive Director")
out <- quiet(split_titles(standardize_conj(remove_dates(standardize_df(example_frame(x))))))
out <- out[order(out$F9_07_COMP_DTK_NAME_PERS, out$Num.Titles), ]
knitr::kable(data.frame(person = out$F9_07_COMP_DTK_NAME_PERS,
                        `before (TitleTxt3)` = out$TitleTxt3,
                        `after (TitleTxt4)` = out$TitleTxt4,
                        Num.Titles = out$Num.Titles, check.names = FALSE), row.names = FALSE)
```

Nine people become `r nrow(out)` rows.

## What it does

```
split_titles()
|
+- apply_misc_split_rules()  fixes for edge cases (PRESIDENT & CEO -> CEO, ...)
+- identify_split_num()      how many titles does the string hold?
+- remove_first_split()      peel off the text before the first "&", repeat
```

Later steps work on one title per row. The person-level view comes back in
[step 08](step-08-categorize-titles.html), which counts each person once
(`title.count`, `title.order`), and [step 09](step-09-conditional-logic.html),
which gives each person one role.

## Assets

Step 04 uses no crosswalk or word list: it relies on the `&` separators that
step 03 wrote.

## Try it

```{r}
identify_split_num("CFO & TREASURER & DIRECTOR")   # 3
remove_first_split("CFO & TREASURER & DIRECTOR")   # " TREASURER & DIRECTOR"
```

```{r}
data(tinypartvii)
d <- dplyr::sample_n(tinypartvii, 120)
df <- d |> standardize_df() |> remove_dates() |> standardize_conj() |> split_titles()
df |>
  dplyr::filter(grepl("&", TitleTxt3)) |>
  dplyr::select(F9_07_COMP_DTK_NAME_PERS, Num.Titles, TitleTxt3, TitleTxt4) |>
  head(8)
```

**Input:** `TitleTxt3`. **Output:** one row per title, with the single title
in `TitleTxt4` and its position within the person's entry in `Num.Titles`.
Continue with [step 05](step-05-standardize-spelling.html).
