---
title: "Step 07 - Standardize titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 07 - Standardize titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
```

**`standardize_titles()`** maps each cleaned title to a canonical `title.standard`
using the title-standardization crosswalk, after applying a few conditional
"executive fixes" for ambiguous finance and chancellor roles.

```
standardize_titles()
|
+- basic_csuite_fixes()   conditional logic -> TitleTxt7
|    +- replace_ceo()     CHANCELLOR -> CEO if paid, else BOARD PRESIDENT
|    +- replace_cfo()     finance roles -> CFO when officer + paid + full-time
+- merge with the crosswalk on the title variant -> title.standard
```

## Crosswalk mapping

The crosswalk collapses many spellings onto one canonical form (for example
"ADMINISTRATION ASSISTANT" and "ADMINISTRATIVE ASSISTANT" both map to
"ADMINISTRATIVE ASSISTANT"). See the [Dictionaries](dictionaries.html) article.

```{r}
library(dplyr)
data(tinypartvii)
set.seed(7)
d <- dplyr::sample_n(tinypartvii, 200)

df <- d |>
  standardize_df() |> remove_dates() |> standardize_conj() |>
  split_titles() |> standardize_spelling() |> gen_status_codes() |>
  standardize_titles()

df |>
  dplyr::select(TitleTxt6, title.standard) |>
  dplyr::distinct() |>
  head(12)
```

## A necessary caveat

The crosswalk is a deterministic 1:1 map, so it cannot resolve genuinely ambiguous
titles — "PRESIDENT" may be a working CEO or an unpaid board president; "DIRECTOR"
may be an executive director or a board member. Step 08 attaches the organization-
level features (pay rank, hours, officer flags) that a later disambiguation step
uses to settle those cases.

**Input:** `TitleTxt6`. **Output:** `TitleTxt7` and the canonical `title.standard`.
Continue with [step 08](step-08-categorize-titles.html).
