---
title: "Step 09 - Resolve roles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 09 - Resolve roles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
```

The crosswalk maps each title to one role, but some titles mean different things
in different organizations:
- A **PRESIDENT** is the paid chief executive in one organization and a volunteer
  board chair in the next.
- A **VICE PRESIDENT** may sit on the board or on staff.
- A **DIRECTOR** may be a board member or a department head.

**`conditional_logic()`** settles these with what the filer reported about each
person: the Part VII checkboxes (trustee, officer, key employee, highly
compensated), pay, and hours. It also names the chief executive when no title
does.

```
conditional_logic()  ->  resolve_roles()
|
+- position, per row       checkboxes first: trustee box = board, officer box = officer,
|                          both = working board; a paid executive title counts as an
|                          officer even without the box; 990-EZ returns use title and pay
+- one role per person     split titles (PRESIDENT/CEO) resolve to one person, one role
+- leadership, per filing  a CEO-level title in an officer position = designated CEO;
                           else the highest-paid plausible leader = imputed CEO;
                           else the filing is board-governed (no CEO is named)
```

## Columns added

| Column | Values |
|---|---|
| `role.final` | CEO, OFFICER, MANAGER, PROFESSIONAL, STAFF, BOARD |
| `role.board` | for BOARD: CHAIR, VICE CHAIR, SECRETARY, TREASURER, MEMBER |
| `role.ceo` | for CEO: designated, co-ceo, interim, imputed |
| `role.position` | board, officer, board_officer, staff, unknown |
| `role.source` | checkbox, title+pay, title |
| `role.primary` | TRUE on one row per person (count people with it) |
| `org.leadership` | per filing: designated, imputed, board_governed, no_paid |

The crosswalk's own `emp.level` and `board.role`, and the legacy flags, are left
as they were: `role.final` is the resolved answer, the crosswalk columns the
title-only default.

```{r}
data(tinypartvii)
set.seed(8)
d <- dplyr::sample_n(tinypartvii, 250)

df <- d |>
  standardize_df() |> remove_dates() |> standardize_conj() |>
  split_titles() |> standardize_spelling() |> gen_status_codes() |>
  standardize_titles() |> categorize_titles() |> conditional_logic()

people <- dplyr::filter(df, role.primary)
table(people$role.final)
table(people$role.board)
```

Titles the crosswalk files as board seats but that resolve otherwise, here a
paid board president who runs the organization:

```{r}
people |>
  dplyr::filter(title.standard == "BOARD PRESIDENT", role.final != "BOARD") |>
  dplyr::select(org.name, title.raw, tot.comp, tot.hours, role.final, role.ceo) |>
  head(8)
```

How each filing is led:

```{r}
df |>
  dplyr::distinct(object.id, formtype, org.leadership) |>
  dplyr::count(formtype, org.leadership)
```

`board_governed` is a valid answer, not a gap: a filing that pays people but
has no plausible paid leader is run by its board. An imputed leader must work
at least 15 hours a week or earn at least $25,000.

**Input:** the output of `categorize_titles()`. **Output:** the same rows with
the role columns added.
