---
title: "Step 08 - Categorize titles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 08 - Categorize titles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
library(titleclassifier)
library(dplyr)
# example titles as a Part VII frame, one person per title
example_frame <- function(titles) {
  d <- titleclassifier::tinypartvii[rep(1, length(titles)), ]
  d$F9_07_COMP_DTK_TITLE <- titles
  d$F9_07_COMP_DTK_NAME_PERS <- sprintf("PERSON %02d", seq_along(titles))
  rownames(d) <- NULL
  d
}
# the pipeline steps print a progress line; keep it out of the tables
quiet <- function(x) { utils::capture.output(x <- force(x)); x }
```

**`categorize_titles()`** attaches the taxonomy to each standardized title,
adds features that describe each person's place in their filing, and renames
the columns to the tidy output schema.

## Before and after

```{r, echo = FALSE}
x <- c("Exec. Dir.", "Chief Operating Officer", "VP Finance and Administration", "Gen. Counsel",
       "Dir of Development", "Director of Nursing", "Mgr, IT", "Head Football Coach", "Staff Physician",
       "Pastor", "Bookkeeper", "Executive Assistant", "Chairman of the Board", "Vice Chair", "Sec",
       "Treasurer", "Director", "Senior Rabbi")
out <- quiet(categorize_titles(standardize_titles(gen_status_codes(standardize_spelling(split_titles(
  standardize_conj(remove_dates(standardize_df(example_frame(x))))))))))
out <- out[order(out$F9_07_COMP_DTK_NAME_PERS), ]
knitr::kable(data.frame(`before (title.standard)` = out$title.standard,
                        domain.category = out$domain.category, domain.label = out$domain.label,
                        SOC = out$detailed.occupation, soc.label = out$soc.label,
                        emp.level = out$emp.level, board.role = out$board.role,
                        check.names = FALSE), row.names = FALSE)
```

Each title is either a job (an `emp.level` and, usually, a SOC occupation) or a
board seat (a `board.role`), never both. An uncoded title (SENIOR RABBI) gets
no taxonomy fields. The [Taxonomy](taxonomy.html) article explains every
category.

## What it does

```
categorize_titles()
|
+- merge with the taxonomy on title.standard   domain, SOC codes, emp.level, board.role
+- add_features()        pay and hours ranks and shares, per-filing counts
+- simplify_varnames()   rename to the output schema (title.v1 ... title.v7, tot.comp, ...)
```

### Taxonomy fields

| Column | Meaning |
|---|---|
| `domain.category`, `domain.label` | the functional area (executive / general management, operations / finance, industry-specific / medical, ...) |
| `major.group` ... `detailed.occupation`, `soc.label` | the 2018 BLS Standard Occupational Classification code, at four levels |
| `emp.level` | for a job: CEO, OFFICER, MANAGER, PROFESSIONAL or STAFF |
| `board.role` | for a board seat: CHAIR, VICE CHAIR, SECRETARY, TREASURER or MEMBER |
| `emp`, `ceo`, `c.level`, `mgr`, `spec`, `board`, `pres`, `vp`, `sec`, `treas`, `mem` | legacy 0/1 flags, derived from `emp.level` and `board.role` |

### Engineered features

`add_features()` computes, within each filing (`object.id`):

* each person's pay and hours: `tot.comp`, `tot.hours` (and versions that
  include related organizations), `pay.rank`, `hours.rank`, `pay.pct.of.max`,
  `pay.pct.of.tot`, `hours.pct.of.max`;
* the filing's make-up: `num.dtk` (people), `num.titles`, `num.paid`,
  `num.fte` (40+ hours), `num.fte.30h`, `num.pte`;
* counts by taxonomy: `num.emp`, `num.board`, `num.ceos`, `num.clevel`,
  `num.mgr`, `num.spec`, `num.pres`, `num.vp`, `num.treas`, `num.sec`,
  `num.mem`;
* a person's titles: `title.count` and `title.order` (from step 04),
  `title.mult.x`.

People with several titles are weighted so that they are not counted twice.

```{r}
data(tinypartvii)
big <- tinypartvii |> group_by(OBJECTID) |>
  filter(RETURN_TYPE == "990", n() >= 12, any(as.numeric(F9_07_COMP_DTK_COMP_ORG) > 150000)) |> ungroup()
one <- tinypartvii[tinypartvii$OBJECTID == big$OBJECTID[1], ]
df <- one |>
  standardize_df() |> remove_dates() |> standardize_conj() |> split_titles() |>
  standardize_spelling() |> gen_status_codes() |> standardize_titles() |> categorize_titles()
df |>
  select(title.standard, emp.level, board.role, tot.comp, tot.hours, pay.rank, num.dtk, num.paid, num.board) |>
  arrange(pay.rank) |>
  head(10)
```

## Assets

| Asset | Type | Used for | Load it with |
|---|---|---|---|
| `title.taxonomy` | crosswalk (package data) | for each `title.standard`: domain, SOC codes, `emp.level`, `board.role`, legacy flags | `get_title_taxonomy()` or `data(title.taxonomy)` |

The source is `data-raw/crosswalks/title-taxonomy.csv` in the source
repository. The domain vocabulary, with a description of each domain, is
`data-raw/crosswalks/domains.csv`.

```{r}
tax <- get_title_taxonomy()
dim(tax)
table(tax$emp.level, useNA = "ifany")
table(tax$board.role[tax$board.role != ""])
tax[tax$title.standard %in% c("CEO", "DIRECTOR OF DEVELOPMENT", "BOARD TREASURER"),
    c("title.standard", "domain.label", "detailed.occupation", "soc.label", "emp.level", "board.role")]
```

**Input:** the output of step 07. **Output:** one row per person and title
with the taxonomy fields, the status flags, the engineered features, and the
title at each step (`title.raw`, `title.v1` ... `title.v7`). Continue with
[step 09](step-09-conditional-logic.html).
