---
title: "End-to-end workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-end workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
```

`titleclassifier` turns the free-text officer and key-employee titles reported on
IRS Form 990 Part VII into a structured, analysis-ready taxonomy. The work is done
by an eight-step pipeline; each step is a single verb you can run on its own, and
each has its own article in this site.

| Step | Function | What it does |
|----:|----------|--------------|
| 01 | `standardize_df()` | Standardize the raw Part VII data frame. |
| 02 | `remove_dates()` | Strip and flag date fragments. |
| 03 | `standardize_conj()` | Normalize conjunctions, prepositions, separators. |
| 04 | `split_titles()` | Split multi-title strings into one row per title. |
| 05 | `standardize_spelling()` | Harmonize spelling, abbreviations, short forms. |
| 06 | `gen_status_codes()` | Detect, flag, and clean status qualifiers. |
| 07 | `standardize_titles()` | Map titles to canonical forms via a crosswalk. |
| 08 | `categorize_titles()` | Assign taxonomy categories and engineer features. |

## Getting the data

Part VII compensation tables are published in the public NCCS efile store. You can
pull one directly by tax year (each file is large, so it is cached on disk):

```{r, eval = FALSE}
library(titleclassifier)

# fetch + read the 2023 Part VII table in one call
d <- get_partvii(2023)

# or split the two steps (download once, re-read from cache)
paths <- fetch_partvii(2023, dest = "efdata")
d     <- read_partvii(paths)
```

For examples the package ships a demo slice, `tinypartvii`, already in the expected
schema.

```{r}
library(titleclassifier)
library(dplyr)

data(tinypartvii)
set.seed(1234)
d <- dplyr::sample_n(tinypartvii, 300)
dim(d)
```

## Running the pipeline

The steps compose with the base pipe. The title-standardization, status-code, and
taxonomy crosswalks are bundled with the package (a pinned snapshot of the source
Google Sheet), so a run needs no network access.

```{r}
df <- d |>
  standardize_df()    |>   # 01
  remove_dates()      |>   # 02
  standardize_conj()  |>   # 03
  split_titles()      |>   # 04
  standardize_spelling() |> # 05
  gen_status_codes()  |>   # 06
  standardize_titles() |>  # 07
  categorize_titles()      # 08

dim(df)
```

## The result

Each row is one standardized title for one person, carrying the raw text, the
successive cleaned versions (`title.v2` … `title.v7`), the canonical
`title.standard`, taxonomy fields, status flags, and organization-level features
(pay/hours ranks, leadership counts).

```{r}
df |>
  select(dtk.name, title.raw, title.standard, domain.category, tot.comp, tot.hours) |>
  arrange(desc(tot.comp)) |>
  head(10)
```

The successive title versions make the cleaning transparent — you can see exactly
what each step changed:

```{r}
df |>
  select(title.raw, title.v3, title.v5, title.v6, title.standard) |>
  distinct() |>
  head(12)
```

## Refreshing the crosswalks

By default the crosswalk loaders read the bundled snapshot. To pull the current
Google-Sheet version instead (and refresh the snapshot in a source checkout), pass
`refresh = TRUE`:

```{r, eval = FALSE}
xwalk <- get_googlesheets_title_xwalk(refresh = TRUE)
```

See the [Dictionaries and crosswalks](dictionaries.html) article for what each
crosswalk contains.
