---
title: "End-to-end workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-end workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE)
```

`titleclassifier` turns the free-text officer and key-employee titles reported on
IRS Form 990 Part VII into a structured, analysis-ready taxonomy. The work is done
by a nine-step pipeline. Each step is a single function you can run on its own,
and each has its own article on this site, with before-and-after examples and the
crosswalks it uses.

| Step | Function | What it does | Assets |
|----:|----------|--------------|--------|
| [01](step-01-standardize-df.html) | `standardize_df()` | Clean the raw table: titles, numeric pay and hours, 0/1 checkboxes. | Part VII schema |
| [02](step-02-remove-dates.html) | `remove_dates()` | Strip and flag dates. | `date.words` |
| [03](step-03-standardize-conj.html) | `standardize_conj()` | One separator (`&`) between titles. | `likely.titles` |
| [04](step-04-split-titles.html) | `split_titles()` | One row per title. | |
| [05](step-05-standardize-spelling.html) | `standardize_spelling()` | Expand abbreviations, fix spelling. | `fix_*` rules, `possible.titles`, `likely.subjects`, hunspell |
| [06](step-06-gen-status-codes.html) | `gen_status_codes()` | Flag and remove FORMER, INTERIM, EX OFFICIO, ... | `status.codes`, `number.words` |
| [07](step-07-standardize-titles.html) | `standardize_titles()` | Map to a standard title. | `title.xwalk`, `title.patterns` |
| [08](step-08-categorize-titles.html) | `categorize_titles()` | Attach the taxonomy; add per-filing features. | `title.taxonomy` |
| [09](step-09-conditional-logic.html) | `conditional_logic()` | Resolve each person's role from checkboxes, pay, hours and title. | decision rules; the role model for 990-EZ |

[Rules and models](classifiers.html) explains how the rule-based steps and the
990-EZ role model differ. [Taxonomy](taxonomy.html) defines every output category,
and [Performance](performance.html) reports coverage and accuracy.

## Getting the data

Part VII compensation tables (`F9-P07-T01-COMPENSATION`) are published in the
public NCCS efile store, one file per tax year. The
[panel990](https://github.com/Nonprofit-Open-Data-Collective/panel990) package
downloads and reads them, caching each file on disk:

```{r, eval = FALSE}
library(panel990)
dl   <- download_tables(years = 2022, tables = "F9-P07-T01-COMPENSATION")
pvii <- read_tables(dl)$tables[[1]]   # 5.07 million rows, 37 columns
```

`titleclassifier` has its own small helpers for the same files:

```{r, eval = FALSE}
library(titleclassifier)

# fetch + read the 2023 Part VII table in one call
d <- get_partvii(2023)

# or split the two steps (download once, re-read from cache)
paths <- fetch_partvii(2023, dest = "efdata")
d     <- read_partvii(paths)
```

For examples the package ships a demo slice, `tinypartvii`, already in the expected
schema.

```{r}
library(titleclassifier)
library(dplyr)

data(tinypartvii)
set.seed(1234)
d <- dplyr::sample_n(tinypartvii, 300)
dim(d)
```

## Running the pipeline

The steps compose with the base pipe. The title-standardization, status-code, and
taxonomy crosswalks are package data, so a run needs no network access.

```{r}
df <- d |>
  standardize_df()    |>   # 01
  remove_dates()      |>   # 02
  standardize_conj()  |>   # 03
  split_titles()      |>   # 04
  standardize_spelling() |> # 05
  gen_status_codes()  |>   # 06
  standardize_titles() |>  # 07
  categorize_titles() |>   # 08
  conditional_logic()      # 09

dim(df)
```

## The result

Each row is one standardized title for one person, carrying the raw text, the
successive cleaned versions (`title.v2` … `title.v7`), the canonical
`title.standard`, taxonomy fields, status flags, organization-level features
(pay/hours ranks, leadership counts), and the person's resolved role.

```{r}
df |>
  select(title.raw, title.standard, domain.label, tot.comp, tot.hours, role.final, role.ceo) |>
  arrange(desc(tot.comp)) |>
  head(10)
```

`role.primary` marks one row per person, so people (not titles) are counted
with it:

```{r}
people <- filter(df, role.primary)
count(people, formtype, role.final)
```

The successive title versions make the cleaning transparent — you can see exactly
what each step changed:

```{r}
df |>
  select(title.raw, title.v3, title.v5, title.v6, title.standard) |>
  distinct() |>
  head(12)
```

## Editing the crosswalks

The crosswalks are package data built from the CSV tables in
`data-raw/crosswalks/`. To change one, edit its table in a source checkout and
rebuild:

```{r, eval = FALSE}
# from the package root
source("data-raw/crosswalks/build-crosswalks.R")
xwalk <- get_title_xwalk()
```

See the [Dictionaries and crosswalks](dictionaries.html) article for what each
crosswalk contains.
