---
title: "Performance"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE, echo = FALSE)
# the tables come from the validation scripts in data-raw/partvii-validation/
RES <- file.path("..", "..", "data-raw", "partvii-validation", "panel-results")
rd  <- function(f) utils::read.csv(file.path(RES, f), check.names = FALSE, stringsAsFactors = FALSE)
pct <- function(x) ifelse(is.na(x), "", paste0(formatC(x, format = "f", digits = 1), "%"))
big <- function(x) ifelse(is.na(x), "", format(x, big.mark = ",", trim = TRUE))
kt  <- function(d, ...) knitr::kable(d, row.names = FALSE, align = c("l", rep("r", ncol(d) - 1)), ...)
size_order <- c("$0", "$1-150k", "$150k-1m", "$1m-10m", "$10m-100m", "$100m+", "missing")
cov  <- rd("perf-coverage.csv")
gold <- rd("perf-gold.csv")
```

How much of Part VII does `titleclassifier` code, and how often is it right?
This page reports two kinds of evidence:

* **Coverage**: the share of titles that reach a standard title, measured on
  a large stratified sample of recent filings.
* **Accuracy**: agreement with hand labels, broken out by form, role, pay and
  organization size. It is backed by checks that need no labels:
  agreement with the filer's own checkboxes, and stability over time.

Every number comes from the scripts in `data-raw/partvii-validation/` of the
source repository, run on the NCCS efile tables (release `efile_v2_3`).

## The data

| Data set | What it is | Used for |
|---|---|---|
| **Calibration sample** | 60,520 filings from 2019-2023: 6,000 organizations followed for all five years, plus 6,000 fresh organizations a year. Stratified by form (990, 990-EZ), size (total expenses, Part I) and number of paid people, with design weights | coverage; checkbox agreement; stability over time |
| **Hand-labeled set** | 316 people on 2010-2012 filings (276 on full 990s, 40 on 990-EZs), drawn to over-represent hard cases: paid presidents, board-run organizations, managers. Each was labeled from the whole filing in two independent passes, with disagreements settled by hand | accuracy |
| **Form switchers** | 23,931 people at organizations that filed a 990-EZ one year and a full 990 the next (or the reverse), matched across years with [synthid](https://github.com/Nonprofit-Open-Data-Collective/synthid). Their full-990 role, decided with the checkboxes, labels their 990-EZ year | accuracy on 990-EZ returns |

Size bins are total expenses: $0, $1-150k, $150k-1m, $1m-10m, $10m-100m and
$100m+. Filers reporting $0 get their own bin because they almost never pay
anyone.

## Coverage

A title row is **coded** when step 07 finds its standard title, by exact match
in the crosswalk or by a pattern rule. It is **uncoded** when neither applies.
A few coded titles are not job titles at all (blank, "SEE SCHEDULE O").

```{r}
ov <- cov[cov$group == "all (weighted)", ]
```

Weighted to the population of 2019-2023 filers, **`r pct(ov$exact + ov$pattern)`
of title rows are coded** (`r pct(ov$exact)` exactly and `r pct(ov$pattern)` by
pattern) and **`r pct(ov$uncoded)` remain uncoded**. Every person still gets a
role: step 09 resolves an uncoded title from the checkboxes, pay and hours.

```{r}
d <- cov
d$group <- c("All filings (sample)", "All filings (weighted)", "990-EZ", "990")[match(d$group, c("all (sample)", "all (weighted)", "990EZ", "990"))]
kt(data.frame(Filings = d$group, `Title rows` = big(d$title_rows), Exact = pct(d$exact),
              Pattern = pct(d$pattern), `Non-job title` = pct(d$non_job_title),
              Uncoded = pct(d$uncoded), `People with a role` = pct(d$people_with_role),
              check.names = FALSE))
```

By size, coverage is highest for small and mid-sized organizations and lowest
for the largest. Their Part VII lists many specialized staff, and their titles
often include unit or system names ("HOSPITAL GROUP RETURN SCHE", "MEMBER AND
PHYSICIAN AND NGPG"):

```{r}
s <- rd("perf-coverage-by-size.csv")
s <- s[order(s$form != "990", match(s$size, size_order)), ]
kt(data.frame(Form = s$form, Size = s$size, `Title rows` = big(s$title_rows),
              Coded = pct(s$exact + s$pattern), Uncoded = pct(s$uncoded), check.names = FALSE))
```

Uncoded titles are a long tail. The `r big(20723)` distinct uncoded titles in
the sample are mostly one-offs: the 25 most common account for only 2.3% of
uncoded rows. The most common are truncations, acronyms and organization
names ("TAF", "AHF", "CLASS A POS", "DIRECTOR CATALIGHT"), which a crosswalk
entry would fix one at a time.

```{r}
u <- head(rd("perf-uncoded-top.csv"), 10)
kt(data.frame(`Uncoded title` = u$title, Rows = u$rows, check.names = FALSE))
```

## Accuracy

Accuracy is the share of hand-labeled people whose resolved role
(`role.final`) matches the label. A person labeled DUAL (a board seat and a
paid executive job at once) counts as correct when resolved as CEO, OFFICER
or BOARD.

```{r}
a <- gold[gold$breakdown == "all", ]
```

**Overall, `r pct(a$accuracy)` of `r a$people` labeled people get the right
role.** Board members on the right seat: `r pct(rd("perf-gold-board-role.csv")$right_board_role)`
of correctly identified board members also get the right board role (chair,
vice chair, secretary, treasurer or member).

### By form

```{r}
g <- gold[gold$breakdown == "form", ]
kt(data.frame(Form = g$group, People = g$people, Accuracy = pct(g$accuracy),
              Method = c("rules, with checkboxes", "model, no checkboxes")))
```

On 990-EZ returns the model replaced rules that scored 82.5% on the same 40
people. On 23,931 form switchers, the model agrees with the full-990-year role
more often than the rules do:

```{r}
sw <- rd("switch-scores.csv")
kt(data.frame(`990-EZ people` = c("All matched", "Same title both years", "... not on the board", "... paid in the 990-EZ year"),
              People = big(sw$people), Rules = pct(sw$step09), Model = pct(sw$model), check.names = FALSE))
```

These scores come from the first version of the model, before the switchers
were added to its training data. The shipped model, scored on the half of the
switchers it did not train on, agrees 96.7% (rules 95.1%) on people with the
same title in both years, and 74.0% (rules 63.6%) on those not on the board.

### By role

Recall is the share of people with a given label who get that role. Precision
is the share of people given a role whose label matches it.

```{r}
r <- gold[gold$breakdown == "labeled role", ]
p <- rd("perf-gold-precision.csv")
m <- data.frame(Role = r$group, Labeled = r$people, Recall = pct(r$accuracy))
m$Resolved <- p$predicted[match(m$Role, p$role)]
m$Precision <- pct(p$precision[match(m$Role, p$role)])
m$Resolved[is.na(m$Resolved)] <- ""
kt(m)
```

Board members, who are most of the people on Part VII, and chief executives are
identified reliably. The hard boundaries are below the CEO: officers versus
managers versus staff. Those lines depend on duties that a title, pay and
hours only partly reveal. STAFF is the weakest role: few people on Part VII
are rank-and-file staff, and step 09 assigns STAFF to more people than are
labeled so.

### By pay

```{r}
g <- gold[gold$breakdown == "pay", ]
kt(data.frame(`Paid by the organization` = g$group, People = g$people, Accuracy = pct(g$accuracy), check.names = FALSE))
```

Unpaid people are almost all board members, and nearly always classified
correctly. Paid people carry the hard cases: is a paid president the CEO,
is a paid director a board member or a department head?

### By size

```{r}
g <- gold[gold$breakdown == "size", ]
kt(data.frame(Size = g$group, People = g$people, Accuracy = pct(g$accuracy)))
```

The labeled set has no $0 filers and only 10 people at $100m+ organizations,
so the size breakdown is indicative only. The checks below cover every size
at scale.

## Checks at scale

These checks use the calibration sample and need no labels. They measure
consistency: they flag problems, but agreement does not prove the answer is
right.

### Agreement with the filer's checkboxes

On a full 990 the filer ticks a box for each person's position. A board title
should come with the trustee box, and a paid executive title with the officer
box:

```{r}
c2 <- rd("calibration-by-size.csv")
c2 <- c2[c2$return_type == "990", ]
c2 <- c2[order(match(c2$size, size_order)), ]
kt(data.frame(Size = c2$size, People = big(c2$people),
              `Board title has trustee box` = pct(c2$board_title_has_trustee_box),
              `Trustee box has board title` = pct(c2$trustee_box_has_board_title),
              `Paid executive title has officer box` = pct(c2$paid_exec_title_has_officer_box),
              `Step 09 changes the title's role` = pct(c2$step09_overrides_title),
              check.names = FALSE))
```

Nearly every trustee box goes with a board title (92-96% at every size). The
reverse is weaker at small organizations: at $1-150k only 77% of board titles
carry the trustee box, because board officers there often tick only the
officer box. At $100m+ only 53% of paid executive titles carry the officer
box: large organizations list many vice presidents as key employees rather
than corporate officers. Step 09 treats a paid executive title as an officer
position with or without the box.

### Finding the chief executive

The share of filings that pay someone and in which step 09 names a CEO, and
how the leadership was found:

```{r}
c3 <- rd("calibration-by-size.csv")
c3 <- c3[order(c3$return_type != "990", match(c3$size, size_order)), ]
kt(data.frame(Form = c3$return_type, Size = c3$size, Filings = big(c3$filings),
              `CEO named, paying filings` = pct(c3$ceo_if_paid),
              Designated = pct(c3$designated), Imputed = pct(c3$imputed),
              `Board-governed` = pct(c3$board_governed), `No one paid` = pct(c3$no_paid),
              check.names = FALSE))
```

Among full-990 filings that pay anyone, a CEO is named in 64% at $1-150k and
over 95% from $1m up. The rest are `board_governed`: small organizations
whose only paid people are a bookkeeper or a part-time helper, or whose paid
president works a few hours for a stipend.

### Stability over time

For the 6,000 organizations followed for five years, people were matched from
one year to the next with synthid. A person's role should rarely change from
one year to the next, and neither should the CEO:

```{r}
st <- rd("crosstime-role-stability.csv")
ce <- rd("crosstime-ceo.csv")
st <- merge(st, ce, by = c("return_type", "size"), all = TRUE)
st <- st[order(st$return_type != "990", match(st$size, size_order)), ]
kt(data.frame(Form = st$return_type, Size = st$size, `Person-years` = big(st$pairs),
              `Same role` = pct(st$same_role), `Org-years` = big(st$org_year_pairs),
              `Same CEO` = pct(st$same_ceo), check.names = FALSE))
```

```{r}
br <- rd("crosstime-by-role.csv")
kt(data.frame(`Role in year 1` = br$role0, `Person-years` = big(br$pairs), `Same role in year 2` = pct(br$same_role),
              check.names = FALSE))
```

95-98% of people keep their role from one year to the next, and in most
groups 88-92% of organizations keep the same CEO. Some change is real (CEOs turn over, board
members become officers), so these rates are a floor on consistency, not an
error rate.

## Reproducing these numbers

The scripts run in order; each states its inputs at the top.

| Script | Produces |
|---|---|
| `12-panel-sample.R`, `13-panel-run.R` | the calibration sample and its classification |
| `14-panel-calibrate.R`, `15-panel-synthid.R` | checkbox agreement, CEO coverage, stability over time |
| `09-gold-check.R` | accuracy on the hand-labeled set |
| `18-switch-labels.R`, `20-role-model-ship.R` | the form-switcher test set and the shipped model |
| `21-performance.R` | the coverage tables and the accuracy breakdowns on this page |
